---
title: "chloro"
output: html_document
---
This R code is for calibrating the CTD fluorometer.

1. Read in the data from past years using the RODBC package
```{r}

###### HERE IS THE BEGINNING OF THE NEW CODE ###########
# Chlorophyll Calibration Script (translated from Steve's SAS code)

# get calibration data from previous cruises
library(RODBC)

# get data through 2012 from MS Access (juv_cruise backup Access database)
# DRIVE LETTER AND/OR PATH MAY VARY
#*** NOTE ***#: HAD TO ADD "dbo_" to table names to make Access query work
channel <- odbcConnectAccess2007("C:/cruise/MWT/juv_cruise_backup24OCT13.mdb",uid="", pwd="")
cal.old.df <- sqlQuery(channel, "SELECT   'C'+CRUISE AS CRUISE_NO,
                                          CTD_INDEX AS INDEX,
                                          BOTTLE_DEPTH AS DEPTH,
                                          CHLOROPHYLL_A AS CHLOR,
                                          FLUOR_VOLT AS VOLTAGE
                                 FROM     dbo_CTD_FLUOR")

# get header data from at_sea database (NOTE LAME SYNTAX REQUIRED BY MS ACCESS)
header.old.df <- sqlQuery(channel, "SELECT 'C'+CRUISE AS CRUISE_NO,
                                           YEAR(CTD_DATE) AS YEAR,
                                           CTD_INDEX AS INDEX,
                                           CTD_NO AS CTD,
                                           DATEPART('h', CTD_DATE) AS HOUR,
                                           DATEPART('n', CTD_DATE) AS MIN
                                    FROM   dbo_CTD_HEADER
                                    WHERE  (CTD_NO IS NOT NULL)")
```

1. Create a simple "day/night"" (D/N) variable defining "day" as hours
between 6am and 9pm.

```{r}
# add "PERIOD" (day/night) to header data
int.hr.old <- with(header.old.df, round(HOUR+(MIN/60),1))
period.old <- ifelse(int.hr.old>6 & int.hr.old<21, "D", "N") 
header.old.df <- cbind.data.frame(header.old.df,
                                  PERIOD = period.old)
rm(int.hr.old, period.old)

# close connection to at_sea database
odbcClose(channel)
rm(channel)

```

1. Read in the data from the most recent year (separate database)

```{r}
# get data through from MS Access (juv_cruise backup Access database)
# DRIVE LETTER AND/OR PATH MAY VARY
channel <- odbcConnectAccess2007("C:/cruise/MWT/cruise_1305/database/at_sea.mdb",uid="", pwd="")
cal.new.df <- sqlQuery(channel, "SELECT   'C'+CRUISE AS CRUISE_NO,
                                          CTD_INDEX AS INDEX,
                                          BOTTLE_DEPTH AS DEPTH,
                                          CHLOROPHYLL_A AS CHLOR,
                                          FLUOR_VOLT AS VOLTAGE
                                 FROM     CTD_FLUOR")

# get header data from at_sea database (NOTE LAME SYNTAX REQUIRED BY MS ACCESS)
header.new.df <- sqlQuery(channel, "SELECT 'C'+CRUISE AS CRUISE_NO,
                                           YEAR(CTD_DATE) AS YEAR,
                                           CTD_INDEX AS INDEX,
                                           CTD_NO AS CTD,
                                           DATEPART('h', CTD_DATE) AS HOUR,
                                           DATEPART('n', CTD_DATE) AS MIN
                                    FROM   CTD_HEADER
                                    WHERE  (CTD_NO IS NOT NULL)")

```

1. Create "day/night" variable for recent data

```{r}
# add "PERIOD" (day/night) to header data
int.hr.new <- with(header.new.df, round(HOUR+(MIN/60),1))
period.new <- ifelse(int.hr.new>6 & int.hr.new<21, "D", "N") 
header.new.df <- cbind.data.frame(header.new.df,
                                  PERIOD = period.new)
rm(int.hr.new, period.new)

# close connection to at_sea database
odbcClose(channel)
rm(channel)


```

1. Merge the old and new data sets into a single dataframe

```{r}
# combine old and new calibration data into single data frame
cal.df <- rbind.data.frame(cal.old.df, cal.new.df)

# combine old and new header data into single data frame
header.df <- rbind.data.frame(header.old.df, header.new.df)

# merge cal.df and header.df on shared columns (CRUISE_NO and INDEX)
# and add column for year/CTD combination ('ID') and log(depth), logz
calibrate.df <- merge(x=cal.df, y=header.df)
calibrate.df <- cbind.data.frame(calibrate.df,
                                 id=paste(calibrate.df[,"YEAR"],"ctd",calibrate.df[,"CTD"], sep=""),
                                 logz=log(calibrate.df$DEPTH))

# convert "id" to a factor
calibrate.df[["id"]] <- as.factor(calibrate.df[["id"]])

```

1. Exclude data from 2007, CTD #5 (collected using Brian Well's methods because carousel was lost that year)

```{r}
calibrate.df <- subset(calibrate.df, id!="2007ctd5")

```

1. Change the default "baseline" factor level to match SAS output

```{r}
# fit GLMs with SAS contrasts to compare with Steve's output
options(contrasts = c("contr.SAS", "contr.poly"))

```

1. Replicate Steve's approach of fitting a linear model and removing data with residuals greater than 2 standard deviations from the mean

```{r}
# fit LM to identify outliers and generate "clean" data
library(MASS)
temp.lm <- lm(CHLOR~id+VOLTAGE+logz, data=calibrate.df)
keepers <- abs(stdres(temp.lm))<=2
clean.df <- calibrate.df[keepers,]

```

1. Fit linear model to 'clean' data

```{r}
# final LM fit
clean.lm <- lm(CHLOR~id+VOLTAGE+logz, data=clean.df)
summary(clean.lm)
anova(clean.lm, test="F")

```

1. Print the equation that predicts chlorophyl for each CTD/year from voltage and depth

```{r}
# Print the equation for 2012 (only one ctd this year)
paste("CHLOR = ",round(coef(clean.lm)["(Intercept)"],4)," + ",
      round(coef(clean.lm)["VOLTAGE"],4),"*VOLTAGE + ",
      round(coef(clean.lm)["logz"],4),"*LOGZ", sep="")

```

1. These plots suggest that we might explore alternative models in the future

```{r}
# plot chlor as a function of voltage, by year/ctd
library(lattice)
xyplot(CHLOR ~ VOLTAGE | id, data=clean.df, type = c("p","r"), col=1)

# plot chlor as a function of logz, by year/ctd
windows()
xyplot(CHLOR ~ logz | id, data=clean.df, type = c("p","r"), col=1)

# FOR FUTURE:
# INVESTIGATE LOG-TRANSFORMATION OF CHLOR AND VOLTAGE
# plot(lm(log(CHLOR)~id+log(VOLTAGE)+logz, data=subset(clean.df, CHLOR>0)))

```

